Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU

### 2016-02-25 paper/paperSource.tex
automatic triggered changes

\emph{\textbf{Data sets.}} In our experiments, six data sets from
different simulations were used. All simulations were done on a
POPC~\footnote{POPC is a chemical compound composed of a
diacylglycerol and phospholipid. Its full name is
1-palmitoyl-2-oleoyl-sn-glycero-3-phosphocholine and it is one of
the most important lipids in bio-physical molecular simulation.}
lipid bilayer, but were all set to produce data of different sizes
(i.e., different number of particles in the simulation). Namely, we
have tested the system on simulations with $52,400$; $209,600$;
$838,400$; $2.5M$; $4.4M$; and $8.8M$ atoms. Also, since the
simulations were run separately, they produced six different MS
systems with distinct characteristics (distinct structure, atom's
positioning, etc.). From all of the generated data sets we have
randomly selected sets of $10$, $100$, and $1000$ consecutive frames
for the purpose of our experiments. This gave us $18$ different
datasets on which we tested our system.

\emph{\textbf{Query work load.}} Two types of query workload were
used: 1) one involving one-body queries only 2) one including
two-body queries (SDH and RDF) as well. The reason for this is that
GROMACS, the system we used to compare our system to, only has a
naive method of solving the RDF (SDH) problem (like almost all MS
analysis systems). In our system we have incorporated SDH (RDF)
algorithms that are far more superior to the naive method, and
comparing the systems like that would not have been fair (we
believe).

\emph{\textbf{One-body queries only.}} The following set of one-body
queries were included in the test workload: mean square displacement
(msd), radius of gyration, dipole moment, center of mass, velocity
autocorrelation, electron density, mass density, and charge density.
This set of queries were pointed to us, by a group in the physics
field with extensive MS background, as one of the most commonly used
in the field of collagen bilayer MS system analysis. A workload
group contains all $8$ queries executed on one of the $12$
selections, making $12$ groups. Such groups are executed on six
different size data sets, with $10$, $100$ and $1000$ frames. This
workload is then repeated $5$ more times, by executing each of the
queries in the groups $5$, $10$, $25$, $50$, and $100$ times,
essentially just magnifying the workload intensity. In total, we
have $12 x 6 x 3 x 6 = 1,296$ different workload setups to test the
system on. Fig.~\ref{fg:workload_set_up} shows the organization of
the workload setup.

@@

\subsection{Benchmark}

Through extensive collaboration with a research group from the
Physics department at USF, we have come up with a benchmark that can
be used for testing the efficiency of an analysis system for
molecular simulations. The benchmark consist of three essential
parts: 1. Simulation data produced by an MS, 2. Queries that are to
be executed onto that data in order to produce some information of
interest, and 3.Benchmark parameters that control the size of the
benchmark.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup10frames52K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results2-10frames}
\end{figure}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup100frames52K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results2-100frames}
\end{figure}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup10frames209K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results3-10frames}
\end{figure}

\begin{figure}
 \centerline{ \includegraphics[width=columnwidth]{images/speedup100frames209K.eps} }
 \caption{Showing speedup for different workload}
 \label{fg:results3-100frames}
\end{figure}

\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5

%\begin{figure}
% \centerline{ \includegraphics[width=.8\columnwidth]{images/speedup10frames4.2M.eps} }
% \caption{Showing speedup for different workload}
% \label{fg:results5-10frames}
%\end{figure}

>>>>

\emph{\textbf{Data sets.}} As it has already been mentioned, we have developed a small python script that generated data sets that are used for the experiments. It accepts two numbers as arguments (number of atoms and frames respectively) and generate a file with appropriate number of atoms/frames and format. We have generated number of mock data sets in range of $1K-8M$ atoms with $10, 100, 1000$ frames each. For each data set we run workloads with the values: $1, 5, 25, 50, 100$ . Number of atoms and frames are obviously up to the scientist and depend on simulations, but workloads are a bit trickier. First of all, they give us a bit better understanding of the average value, which might vary on different circumstances related to OS, secondly, once this system is integrated into a distributed system, we can imagine that each frame could be processed by a separate node keeping the values inside of the appropriate memory (global or RAM) and generate responses based on the clients selections requests. This possibility will be elaborated in more details in Future Work section.

\emph{\textbf{SDH implementation.}} Even though implementation of SDH can be different, in the previous paper the algorithm in the previous paper was based on quad tree data structure with data proximity approach based on data uniformity\cite{mainPaper}, but in our experiments we follow special CUDA based approach which perfectly fits the nature of GPU devices with the given memory hierarchy, and the number of computations are equal to brute force, consequently, there is no expectation on data uniformity, thus, the problem solution complexity is still the same with the worst case of $O(N^2)$, but processed in parallel by GPU cores.

The following set of one-body
queries were included in the test workload: mean square displacement
(msd), radius of gyration, dipole moment, center of mass, velocity
autocorrelation, electron density, mass density, and charge density.
This set of queries were pointed to us, by a group in the physics
field with extensive MS background, as one of the most commonly used
in the field of collagen bilayer MS system analysis. A workload
group contains all $8$ queries executed on one of the $12$
selections, making $12$ groups. Such groups are executed on six
different size data sets, with $10$, $100$ and $1000$ frames. This
workload is then repeated $5$ more times, by executing each of the
queries in the groups $5$, $10$, $25$, $50$, and $100$ times,
essentially just magnifying the workload intensity. In total, we
have $12 x 6 x 3 x 6 = 1,296$ different workload setups to test the
system on. Fig.~\ref{fg:workload_set_up} shows the organization of
the workload setup.

\begin{figure}
 \centerline{ \includegraphics[width=1.0\columnwidth]{images/chart_1000}}
 \caption{Speedup for 1000 atoms}
 \label{fg:chart_1000}
\end{figure}

\begin{figure}
 \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_10000}}
 \caption{Speedup for 10000 atoms}
 \label{fg:chart_10000}
\end{figure}

\begin{figure}
 \centerline{\includegraphics[width=1.0\columnwidth]{images/chart_max}}
 \caption{Speedup for larger number of atoms}
 \label{fg:chart_max}
\end{figure}

@@

\subsection{Benchmark}
Through extensive collaboration with a research group from the
Physics department at USF, we have come up with a benchmark that can
be used for testing the efficiency of an analysis system for
molecular simulations. The benchmark consist of three essential
parts: 1. Simulation data produced by an MS, 2. Queries that are to
be executed onto that data in order to produce some information of
interest, and 3.Benchmark parameters that control the size of the
benchmark.

@@
